---
title: "Functional Programming with Purrr"
author: "Juan Carlos Villaseñor"
date: "13 de enero de 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)

suppressPackageStartupMessages({
  library(tidyverse)
})


```


## Ex. 1

```{r}

load("data/swapi.rda")

length(people)

people[[1]]

```


## Mapping function

`map(.x, .f)` = for each element in x, do f

- Takes two arguments, x and f
- x is a vector, list, or dataframe
- f... a function?

Strategy is often to find out what to do for one element. Here, count the number of starships for Luke.

```{r}

luke <- people[[1]]

n.luke.starships <- length(luke$starships)

```

By doing `~length(.x$starships)` we can convert the above logic into a function.

```{r}

map(people, ~length(.x$starships))

```


### Task: Find the name of each character's homewoeld

First do it for Luke

```{r}

load("data/planet_lookup.rda")

luke.homeworld <- luke$homeworld

planet_lookup[luke.homeworld]

```

We can simplify for Luke as

```{r}
luke.homeworld <- planet_lookup[luke$homeworld]
```

And we can generalize it using map

```{r}
map(people, ~planet_lookup[.x$homeworld])
```

## ROAD`map()`

Additional functions include:

  - `map_lgl()` will return a logical vector
  
  - `map_int()` will return an integer vector
  
  - `map_dbl()` will return a double
  
  - `map_chr()` will return a character
  


```{r}

# We rename the lists
people <- people %>% set_names(map_chr(people, "name"))

# How many starships has each character been in?

map_int(people, ~length(.x[["starships"]]))

# What is the hair color of each character?

map_chr(people, ~.x[["hair_color"]])

# Which characters are males?

map_lgl(people, ~.x[["gender"]] == "male")

# How heavy people are?

map_chr(people, ~.x[["mass"]]) #%>% #Some info is unknown, so we better stick to character
  #readr::parse_number(na = "unknown")

```

  

## Some side-notes on `map(.x, .f)`

- `.f` can be a string or an integer

- `map_chr(.x = people, .f = "mass")` would have also worked

Along with the pipe operator, we can see how many starships every character has been in

```{r}

map(people, "starships") %>% map_int(length)

```

## Other tasks

Which film as the most characters?

```{r}

map(people, "films") %>%
  map_int(length) %>%
  sort(decreasing = T)

```


Which film has more characters?

```{r}

map(films, "characters") %>% #Extract the characters in each fil
  map_int(length) %>%        # Get the number of character in each film
  set_names(map_chr(films, "title")) %>% #Rename the vectors using the tile of the movie
  sort(decreasing = T)      # Order the results in a decreasing way

```


# `purrr` and list columns

```{r}

people_tbl <- tibble(
  name = people %>% map_chr("name"),
  
  films = people %>% map("films"),
  
  height = people %>% map_chr("height") %>% 
    readr::parse_number(na = "unknown"),
  
  species = people %>% map_chr("species", .null = NA_character_)
)

# The above is equivalent to the following:


people_tbl2 <- tibble(
  name = map_chr(people, "name"),
  
  films = map(people, "films"),
  
  height = map_chr(people, "height") %>% 
    readr::parse_number(na = "unknown"),
  
  species = map_chr(people, "species", .null = NA_character_)
)

```


### Task: Add a column with the number of movies

```{r}

## create some lookup tables

film_lookup <- map_chr(films, "title") %>% set_names(map_chr(films, "url"))
film_number_lookup <- stringr::str_split_fixed(names(film_lookup), "/", 7)[, 6]  %>%
  as.numeric() %>%
  set_names(map_chr(films, "url"))

people_tbl <- people_tbl %>%
  mutate(
    n_films = map_int(films, length),
    film_names = map(films, ~ film_lookup[.x]),
    film_numbers = map(films, ~ film_number_lookup[.x])
  )

```


### New column that collapses film numbers into a single string

```{r}
people_tbl <- people_tbl %>% 
  mutate(
    movies_squashed = map_chr(film_numbers, paste, collapse = ", ")
  )
```























